# Reproducible Research -- Week 2 Assignment

This is a combined R code and Markdown document for knitr. This will generate an HTML output.

```{r}
library( data.table )
library( ggplot2 )
```
## Section A: Loading and preprocessing the data

### Question A.1
```{r load_file}
df <- read.csv( "repdata-data-activity/activity.csv")
```

### Question A.2
Split the data into two tables, one containing all the steps=NA rows and the other without NA

```{r proprocess}
dt <- data.table( df )  ## use a data table structure for ease of processing

dt_na <- subset(dt, is.na(steps))
dt_notna <- subset(dt, !is.na(steps))
```

## Section B: What is mean total number of steps taken per day?
Now compute the mean and median number of steps per day (where the values are not NA)

```{r mean_byday}
d_byday <- dt_notna[, list(total_steps = sum(steps), mean = mean(steps), median = median(steps)), by=date]
```

### Question B.1
Create a histogram of the total number of steps per day.

```{r total_steps_byday}
qplot( total_steps, data=d_byday )
```

### Question B.3
Display the total, mean and median steps by day.

```{r}
d_byday
```

## Section C: What is the average daily activity pattern?

Similarly, let's calculate the mean/average and maximum number of steps by interval.

```{r mean_byinterval}
d_interval <- dt_notna[, list(interval_mean=mean(steps), interval_max = max(steps)), by=interval]
```
### Question C.1
Let's plot the mean number of steps by interval:
```{r mean_steps_byinterval}
ggplot(d_interval, aes(interval,interval_mean)) + geom_line()
```

For fun, let's also plot the max number of steps by interval:
```{r max_steps_byinterval}
ggplot(d_interval, aes(interval,interval_max)) + geom_line()
```

### Question C.2
To answer the question "Which interval, across all days, has the maximum number of steps,
we do the following:
```{r}
max(d_interval$interval_max)
```
##Section D: Imputing missing values

### Question D.1
There are *`r nrow(dt_na)` rows* with NAs.

### Question D.2
We will use the interval mean to fill in the NA values. However, since the mean
is a non-integer, it needs to be converted to an integer before assigning it to
the steps column.

```{r} 
dt_na$steps <- d_interval[ match(dt_na$interval, d_interval$interval), as.integer(interval_mean) ]
```
### Question D.3
Re-combine the two data tables that had been split by NA and not-NA, then order the resultant table
by (date, interval)

```{r recombine}
dt2 <- dt_na
dt2 <- rbind( dt2, dt_notna )
dt2 <- dt2[order(date,interval),]
```

### Question D.4

```{r total_steps_byday_simulated}
d_byday_notna <- dt2[, list(total_steps = sum(steps), mean_steps = mean(steps), median_steps = median(steps)), by=date]

qplot( total_steps, data=d_byday_notna )
```


## Section E: Are there differences in activity patterns between weekdays and weekends?

### Question E.1
Add a column "dayofweek" to distinguish between weekdays and weekends.

```{r dayofweek}

v <- sapply( dt2$date, function(x) { if (weekdays(as.Date(x), abbreviate=FALSE) %in% c("Saturday","Sunday")) "weekend" else "weekday" })

dt2 <- cbind( dt2, dayofweek = as.factor(v) )
```

### Question E.2

```{r steps_dayofweek_simulated}
qplot( interval, steps, data=dt2, facets=dayofweek~. )
```

